---
title: "Introduction to SCP: Single Cell Pipeline"
author: "Hao Zhang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to SCP: Single Cell Pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)
```

## Overview

**SCP** (Single Cell Pipeline) is a comprehensive R package designed to facilitate end-to-end analysis and exploration of single-cell RNA-seq data. This vignette provides a quick introduction to the main features and workflows of SCP.

### Key Features

- **Quality Control**: Comprehensive cell and feature QC with multiple metrics
- **Data Integration**: 15+ integration methods for batch correction
- **Dimensionality Reduction**: Standard and advanced methods (PCA, UMAP, tSNE, etc.)
- **Cell Annotation**: Automated cell type annotation with multiple algorithms
- **Trajectory Analysis**: Pseudotime and lineage inference
- **Rich Visualization**: 100+ plotting functions for comprehensive data exploration
- **Python Integration**: Seamless integration with Python tools (scVelo, PAGA, etc.)

## Installation

```{r installation, eval=FALSE}
# Install from GitHub
if (!require("devtools")) install.packages("devtools")
devtools::install_github("zhanghao-njmu/SCP")
```

## Quick Start

### Loading the Package

```{r load-package}
library(SCP)
library(Seurat)
```

### Example Data

SCP includes several example datasets for demonstration:

```{r load-data}
# Load example pancreatic data
data("pancreas_sub")

# Examine the object
pancreas_sub
```

This dataset contains `r ncol(pancreas_sub)` cells and `r nrow(pancreas_sub)` genes from mouse pancreatic endocrinogenesis.

### Basic Workflow

#### 1. Quality Control

The first step in single-cell analysis is quality control:

```{r qc, eval=FALSE}
# Run cell QC
pancreas_sub <- RunCellQC(
  srt = pancreas_sub,
  db_method = "scDblFinder",  # Doublet detection method
  return_filtered = FALSE      # Return object with QC metrics
)

# Visualize QC metrics
CellStatPlot(
  srt = pancreas_sub,
  stat.by = c("nFeature_RNA", "nCount_RNA", "percent_mito"),
  group.by = "CellQC"
)
```

#### 2. Data Normalization and Integration

For datasets with multiple batches:

```{r integration, eval=FALSE}
# Standard workflow including normalization and integration
pancreas_sub <- Standard_SCP(
  srt = pancreas_sub,
  batch = "tech",                    # Batch variable
  integration_method = "Harmony",     # Integration method
  normalization_method = "LogNormalize"
)
```

#### 3. Dimensionality Reduction

```{r dim-reduction, eval=FALSE}
# Run multiple dimensionality reduction methods
pancreas_sub <- RunDimReduction(
  srt = pancreas_sub,
  methods = c("PCA", "UMAP", "tSNE"),
  dims = 1:30
)
```

#### 4. Cell Type Annotation

```{r annotation, eval=FALSE}
# Automated cell type annotation
pancreas_sub <- RunSingleR(
  srt = pancreas_sub,
  ref = "ref_scMCA",  # Reference dataset
  species = "Mus_musculus"
)
```

### Visualization

SCP provides extensive visualization capabilities:

#### Dimension Plots

```{r dim-plot, eval=FALSE}
# Cell dimension plot colored by cell type
CellDimPlot(
  srt = pancreas_sub,
  group.by = "celltype",
  reduction = "UMAP",
  theme_use = "theme_scp"
)
```

#### Feature Plots

```{r feature-plot, eval=FALSE}
# Expression of specific genes
FeatureDimPlot(
  srt = pancreas_sub,
  features = c("Ins1", "Gcg", "Sst"),
  reduction = "UMAP",
  theme_use = "theme_scp"
)
```

#### Heatmaps

```{r heatmap, eval=FALSE}
# Group heatmap showing marker genes
GroupHeatmap(
  srt = pancreas_sub,
  features = c("Ins1", "Gcg", "Sst", "Ppy"),
  group.by = "celltype",
  show_row_names = TRUE
)
```

### Differential Expression Analysis

```{r de-analysis, eval=FALSE}
# Find markers between cell types
markers <- RunDEtest(
  srt = pancreas_sub,
  group_by = "celltype",
  fc.threshold = 1.5,
  only.pos = TRUE
)

# Visualize with volcano plot
VolcanoPlot(
  srt = pancreas_sub,
  de_result = markers,
  group_by = "celltype"
)
```

### Enrichment Analysis

```{r enrichment, eval=FALSE}
# Run enrichment analysis on marker genes
enrichment <- RunEnrichment(
  srt = pancreas_sub,
  db = "GO_BP",
  species = "Mus_musculus"
)

# Visualize results
EnrichmentPlot(
  enrichment_result = enrichment,
  plot_type = "bar"
)
```

## Advanced Features

### Trajectory Analysis

```{r trajectory, eval=FALSE}
# Trajectory inference with Monocle3
pancreas_sub <- RunMonocle3(
  srt = pancreas_sub,
  group.by = "celltype"
)

# Visualize lineage
LineagePlot(
  srt = pancreas_sub,
  reduction = "UMAP"
)
```

### Python Integration

```{r python-integration, eval=FALSE}
# Prepare Python environment
PrepareEnv(version = "3.8-1")

# Run RNA velocity with scVelo
pancreas_sub <- RunSCVELO(
  srt = pancreas_sub,
  group_by = "celltype"
)

# Visualize velocity
VelocityPlot(
  srt = pancreas_sub,
  reduction = "UMAP"
)
```

### Interactive Exploration

```{r interactive, eval=FALSE}
# Launch interactive Shiny app
RunSCExplorer(
  srt = pancreas_sub,
  browser = TRUE
)
```

## Integration Methods

SCP supports multiple integration methods:

- Harmony
- Seurat (CCA/RPCA)
- fastMNN
- scVI
- LIGER
- ComBat
- BBKNN
- Scanorama
- MNN
- CSS
- And more...

```{r integration-comparison, eval=FALSE}
# Compare different integration methods
pancreas_integrated <- Integration_SCP(
  srtMerge = pancreas_sub,
  batch = "tech",
  integration_method = c("Harmony", "Seurat", "fastMNN")
)
```

## Session Info

```{r session-info}
sessionInfo()
```

## Further Reading

- [Package website](https://zhanghao-njmu.github.io/SCP/)
- [GitHub repository](https://github.com/zhanghao-njmu/SCP)
- Function documentation: `?FunctionName`

For specific analysis workflows, please refer to the function documentation and examples.

## Citation

If you use SCP in your research, please cite:

```
Zhang, H. (2024). SCP: Single Cell Pipeline. R package version 0.5.6.
https://github.com/zhanghao-njmu/SCP
```
